---
title: "CFTR variant explorer"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: 
      version: 4
      bootswatch: litera
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
```

Variant-specific functional deficits {data-icon="fa-signal"}
===================================== 

## Variant-specific CFTR folding versus conductance deficits dotplot

```{r}
data <- read_csv("bihler_2023_clean.csv")
q <- data %>%
  # filter(variant %in% c("WT-ctrl", "G551D-ctrl", "G551D", "F508del","F508del-ctrl", "I507del")) %>%
  # filter( (band_c_pct_wt_DMSO < 10) | (fsk_pct_wt_DMSO < 10)) %>%
    filter(!(is.na(trikafta_response))) %>%
  ggplot(aes(band_c_pct_wt_DMSO, fsk_pct_wt_DMSO, fill = trikafta_response, text  = variant)) +
  geom_point(shape = 21, colour = "black", size = 5, stroke = 0.5) +
  scale_fill_manual(values=c("red4", "grey70")) +
  # xlim(0,20) + 
  # ylim(0,20) + 
  geom_vline(xintercept = c(10), linetype=2) +
  geom_hline(yintercept = c(10), linetype=2) +
  labs(y = "conductance (mean, % WT)", x = "folding (mean, % WT)") +
  ggtitle("Variant-specific CFTR folding and conductance") +

  theme_minimal()
ggplotly(q)
```


Column {data-width=350}
-----------------------------------------------------------------------

### Variant-specific CFTR folding versus conductance deficits by residue

```{r}

# PLOT 5: FOLDING AND CONDUCTANCE FOR EACH RESNO BARPLOT
# devise ordering variables
data <- read_csv("bihler_2023_clean.csv")
data <- data %>% mutate(ave_fold_cond =(band_c_pct_wt_DMSO*fsk_pct_wt_DMSO/2))
data <- data %>% mutate(rownum = row_number()) 
data <- data %>% unite(ord, c(ave_fold_cond,rownum), sep = "", remove = FALSE)
data$ord <- as.numeric(data$ord)
data <- data %>% unite(variant2, c(variant,rownum), sep = ".", remove = FALSE)
data$variant2 <- factor(data$variant2, levels = data$variant2[order(data$ord)])

# order by variant
p <- data %>%
  ggplot(aes(fill = trikafta_response, color = trikafta_response, text  = variant)) +
  geom_segment(aes(x=variant2, xend=variant2, y=fsk_pct_wt_DMSO, yend=band_c_pct_wt_DMSO)) +
  geom_point( aes(x=variant2, y=fsk_pct_wt_DMSO), size=1 ) +
  geom_point( aes(x=variant2, y=band_c_pct_wt_DMSO), size=1) +
  scale_color_manual(values=c("red4", "grey70")) +
  scale_fill_manual(values=c("red4", "grey70")) +
  theme_minimal()
ggplotly(p)

```

### Variant-specific CFTR folding versus conductance deficits by mean functional deficit

```{r}
data <- read_csv("bihler_2023_clean.csv")

# devise ordering variables
data <- read_csv("bihler_2023_clean.csv")
data <- data %>% mutate(ave_fold_cond =(band_c_pct_wt_DMSO*fsk_pct_wt_DMSO/2))
data <- data %>% mutate(rownum = row_number()) 
data <- data %>% unite(ord, c(ave_fold_cond,rownum), sep = "", remove = FALSE)
data$ord <- as.numeric(data$ord)
data <- data %>% unite(variant2, c(variant,rownum), sep = ".", remove = FALSE)
data$variant2 <- factor(data$variant2, levels = data$variant2[order(data$ord)])

# order by mean fold/conductance
q <- data %>%
    filter(!(is.na(trikafta_response))) %>%
  ggplot(aes(fill = trikafta_response, text  = variant)) +
  geom_bar(aes(variant2, fsk_pct_wt_DMSO), stat = "identity") +
  geom_bar(aes(variant2, (band_c_pct_wt_DMSO)*(-1)), stat = "identity") +
  scale_fill_manual(values=c("red4", "grey70")) +
  geom_hline(yintercept = c(0)) +
  geom_hline(yintercept = c(10), linetype=2) +
  geom_hline(yintercept = c(-10), linetype=2) +
  theme_minimal()
ggplotly(q)
```



Variant-specific modulator responses {data-icon="fa-solid fa-tablets"}
===================================== 

Column {data-width=350}
-----------------------------------------------------------------------

## Variant-specific CFTR correction by ivacaftor versus texacaftor/elexacaftor response dotplot

```{r}
data <- read_csv("bihler_2023_clean.csv")

t <- data %>%
  # filter(variant %in% c("WT-ctrl", "G551D-ctrl", "G551D", "F508del","F508del-ctrl", "I507del")) %>%
  filter(!(is.na(trikafta_response))) %>%
  filter( (band_c_pct_wt_DMSO < 10) | (fsk_pct_wt_DMSO < 10)) %>%
    filter(!(is.na(trikafta_response))) %>%
  ggplot(aes(fsk_pct_wt_ELXTEZ, fsk_plus_iva_pct_wt_DMSO, fill = trikafta_response, text  = variant)) +
  geom_point(shape = 21, colour = "black", size = 5, stroke = 0.5) +
  scale_fill_manual(values=c("red4", "grey70")) +
  # xlim(0,20) + 
  # ylim(0,20) + 
  geom_vline(xintercept = c(10), linetype=2) +
  geom_hline(yintercept = c(10), linetype=2) +
  theme_minimal()
ggplotly(t)

```

Column {data-width=350}
-----------------------------------------------------------------------

## Variant-specific CFTR correction by ivacaftor versus texacaftor/elexacaftor response by residue

```{r}

data <- read_csv("bihler_2023_clean.csv")
data <- data %>% mutate(ave_iva_elxtez =(fsk_plus_iva_pct_wt_DMSO*fsk_pct_wt_ELXTEZ/2))
data <- data %>% mutate(rownum = row_number()) 
data <- data %>% unite(ord, c(ave_iva_elxtez,rownum), sep = "", remove = FALSE)
data$ord <- as.numeric(data$ord)
data <- data %>% unite(variant2, c(variant,rownum), sep = ".", remove = FALSE)
data$variant2 <- factor(data$variant2, levels = data$variant2[order(data$ord)])

# order by resno
r <- data %>%
    filter(!(is.na(trikafta_response))) %>%
  ggplot(aes(fill = trikafta_response, text  = variant)) +
  geom_bar(aes(resno, fsk_plus_iva_pct_wt_DMSO), stat = "identity") +
  geom_bar(aes(resno, (fsk_pct_wt_ELXTEZ)*(-1)), stat = "identity") +
  scale_fill_manual(values=c("red4", "grey70")) +
  geom_hline(yintercept = c(0)) +
  geom_hline(yintercept = c(10), linetype=2) +
  geom_hline(yintercept = c(-10), linetype=2) +
  theme_minimal()
ggplotly(r)

```

## Variant-specific CFTR correction by ivacaftor versus texacaftor/elexacaftor response by residue

```{r}
data <- read_csv("bihler_2023_clean.csv")
data <- data %>% mutate(ave_iva_elxtez =(fsk_plus_iva_pct_wt_DMSO*fsk_pct_wt_ELXTEZ/2))
data <- data %>% mutate(rownum = row_number()) 
data <- data %>% unite(ord, c(ave_iva_elxtez,rownum), sep = "", remove = FALSE)
data$ord <- as.numeric(data$ord)
data <- data %>% unite(variant2, c(variant,rownum), sep = ".", remove = FALSE)
data$variant2 <- factor(data$variant2, levels = data$variant2[order(data$ord)])

s <- data %>%
    filter(!(is.na(trikafta_response))) %>%
  ggplot(aes(fill = trikafta_response, text  = variant)) +
  geom_bar(aes(variant2, fsk_plus_iva_pct_wt_DMSO), stat = "identity") +
  geom_bar(aes(variant2, (fsk_pct_wt_ELXTEZ)*(-1)), stat = "identity") +
  scale_fill_manual(values=c("red4", "grey70")) +
  geom_hline(yintercept = c(0)) +
  geom_hline(yintercept = c(10), linetype=2) +
  geom_hline(yintercept = c(-10), linetype=2) +
  theme_minimal()
ggplotly(s)

```




